Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-10-11 code/rmarkdown/article.Rmd
+ fix refs + use one-sided tests in post hoc for second analysis

<!-- Introduction -->
Protected areas spearhead conservation efforts [@r440]. They buffer species from anthropogenic impacts; providing places for them to persist. Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255]. To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes. Intra-specific genetic variation is fundamental to long-term species persistence [@r165], and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].

Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg. @r500]. By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@r67]. Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation [@r67; @r3; @r43].

Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently intra-specific genetic variation--without needing to directly utilise genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. @r202; @r2]. By generating prioritisations that preserve a more representative sample of environmental conditions across species' ranges, these prioritisations could also capture more adaptive variation. However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified. Neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, isolation by distance [IBD; @r259] predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci. Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. habitat type @r497]--the typical situations in most conservation planning exercises.

Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning. We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451]. We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables. We also show that for most species the level of association is strong enough to be operational for conservation planning. Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation than expected by chance.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The effectiveness of prioritizations at securing genetic variation. Box plots show the proportion of adaptive and neutral genetic variation secured for each species using different targets and under different scenarios. Red lines show the target proportion of genetic variation to secure a representative sample.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>% 
	group_by(Prioritisation, Context, Metric, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic surrogates are strong enough to be operational (Figure 2). For most species, the environmental and geographic surrogates variables were effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57). Despite this, for a few species, geographic distance was a poor surrogate for neutral genetic variation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). Indeed, geographic distances will be a poor surrogate for species that utilise dispersal modes that result in non-linear relationships between connectivity and geographic distance [eg. wind dispersing plants can maintain high connectivity between distant areas; @r451], and in places where connectivity is most affected by landscape features [@r497]. Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well for a diverse set of species.

@@

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Second, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography. However, such distances require specific-specific scaling to adequate describe their connectivity [eg. @r497], and so cannot easily be utilised in large-scale planning exercises. Third, we used amplified fragment length polymorphism [AFLP; @r453] data to characterise the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [@r480], no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project [@r451; Figure 2a]. This data set has previously been used to explore patterns of adaptive [eg. @r475; @r455] and neutral genetic variation [eg. @r478], and to show that species richness is not an effective surrogate for genetic diversity in conservation planning [@r486]. It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in Alvarez _et al._ [@r478] for more information). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell. We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.

@@

We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were not found to contain individuals belonging to the study species (`r nrow(grid.DF)` grid cells used in analysis). To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; @r472] and used this to represent acquisition cost (Figure 2b). All spatial and statistical analysis were conducted in \texttt{R} [version `r paste0(R.version$major, '.', R.version$minor)` @r192].

We created environmental and geographic surrogate variables for each species. To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers [$30^{\prime\prime}$ resolution; @r33], projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable. To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1). We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2). Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species. In these spaces, each planning unit was associated with a single point. Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).

@@

To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive. Following best practices, we used two outlier detection methods to achieve this [an individual- and a population-level method @r495]. The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection [@r480]. The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis. Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral. To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1 @r456]. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species separately. Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives. We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} [version 2.3.4; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations] using the number of lineages previously determined by Alvarez _et al._ [-@r478] and combining replicate runs using \texttt{ClumPP} [version 1.1.2, @r459; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations). We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species, `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466]. We omitted individuals if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).

The second outlier detection method involved fitting PCAs to identify outlier loci [implemented in the \texttt{pcadapt R} package; @r487]. To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis. For each of the 27 species, we first imputed missing data--replacing missing values with the average frequency among conspecifics--ran a PCA, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci). We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection [using the \texttt{qvalue R} package; @r491].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices. We applied non-metric multi-dimensional scaling analyses [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [via the \texttt{cluster R} package; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had multiple samples per species, we averaged the ordinated values belonging to conspecifics in the same grid cell.

@@

We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' geographic range (using amount-based targets) and a sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values--because they secured a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the \texttt{raptr R} package [@r465].

@@

We then fit maximum likelihood population-effects (MLPE) models [@r494]. These models accommodate the structure of dissimilarity matrices using random effects. For each species, we fit a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation. Additionally, if adaptive loci were detected, we also fit a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation. All predictor variables were _z_-transformed prior to model fitting to improve convergence. To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its corresponding null model using $\chi^2$, and applied Bonferroni corrections.

@@

We fit two full generalised linear models (GLM) with logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation using the proportion of environmental variation also secured in the prioritisation. The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using proportion of geographic variation also secured in the prioritisation. Both models also contained a variable indicating the species for which the prioritisations were generated, and an interaction term. They were subject to backwards step-wise term deletion routines to assess term significance. To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value [using the \texttt{pscl R} package; @r492].

Our third aim was to determine whether surrogate-based targets improved prioritisations. As previously mentioned, we generated prioritisations using different combinations of targets and conservation planning scenarios. We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations. Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets. We generated a single solution for each targets/scenario combination, except for the single-species scenario with just amount-based targets for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because this combination had many optimal solutions. We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.

We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links [GLMMs; using the \texttt{lme4 R} package; @r469]. We fit a full model to the proportion of genetic variation secured in a given prioritisation. This model contained categorical variables indicating the combination of targets (amount-only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (single-species, multi-species, or multi-species with cost), the type of genetic variation measured (adaptive or neutral), and all interactions between them. Prioritisations generated for the same species were accomodated using a random intercept term. We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].

>>>>

<!-- Introduction -->
\dropcap{P}rotected areas spearhead conservation efforts ([@r440]). They buffer species from anthropogenic impacts; providing places for them to persist. Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost ([@r255]). To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes. Intra-specific genetic variation is fundamental to long-term species persistence ([@r165]), and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity ([@r67; @r3; @r61; @r499]).

Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral (reviewed in [@r480]). Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg. [@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg. [@r500]). By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]). In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]). Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).

Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently intra-specific genetic variation--without needing to directly utilise genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg. [@r202; @r2]). By generating prioritisations that preserve a more representative sample of environmental conditions across species' ranges, these prioritisations could also capture more adaptive variation. However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified. Neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]). Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, isolation by distance (IBD; [@r259]) predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci. Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network ([@r200]). However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors (eg. habitat; [@r497])--the typical situations in most conservation planning exercises.

Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning. We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps ([@r451]). We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables. We also show that for most species the level of association is strong enough to be operational for conservation planning. Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation than expected by chance.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='Box plots show the proportion of adaptive and neutral genetic variation secured for each species in solution generated from different targets and scenarios. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>% 
	group_by(Prioritisation, Context, Metric, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic surrogates are strong enough to be operational (Figure 2). For most species, the environmental and geographic surrogates variables were effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57). Despite this, for a few species, geographic distance was a poor surrogate for neutral genetic variation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). Indeed, geographic distances will be a poor surrogate for species that utilise dispersal modes that result in non-linear relationships between connectivity and geographic distance [eg. wind dispersing plants can maintain high connectivity between distant areas; @r451], and in landscapes where connectivity is most affected by landscape features ([@r497]). Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well for a diverse set of species.

@@

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Second, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography. However, such distances require specific-specific scaling to adequate describe their connectivity (eg. [@r497]), and so cannot easily be utilised in large-scale planning exercises. Third, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterise the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a). This data set has previously been used to explore patterns of adaptive (eg. [@r475; @r455]) and neutral genetic variation (eg. [@r478]), and to show that species richness is not an effective surrogate for genetic diversity in conservation planning ([@r486]). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell. We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.

@@

We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were not found to contain individuals belonging to the study species (`r nrow(grid.DF)` grid cells used in analysis). To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b). All spatial and statistical analysis were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).

We created environmental and geographic surrogate variables for each species. To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable. To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1). We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2). Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species. In these spaces, each planning unit was associated with a single point. Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).

@@

To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive. Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]). The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]). The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis. Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral. To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} (version 2.1; [@r456]). We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately. Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives. We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations). We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]). We omitted individuals if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).

The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r487]). To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis. For each of the 27 species, we first imputed missing data--replacing missing values with the average frequency among conspecifics--ran a PCA, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci). We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices. We applied NMDS (implemented in the \texttt{vegan R} package; [@r458]) using Gower distances (via the \texttt{cluster R} package; [@r457]) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had multiple samples per species, we averaged the ordinated values belonging to conspecifics in the same grid cell.

@@

We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection ([@r465]). This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' geographic range (using amount-based targets) and a sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values--because they secured a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).

@@

We then fit maximum likelihood population-effects (MLPE) models ([@r494]). These models accommodate the structure of dissimilarity matrices using random effects. For each species, we fit a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation. Additionally, if adaptive loci were detected, we also fit a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation. All predictor variables were _z_-transformed prior to model fitting to improve convergence. To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its corresponding null model using $\chi^2$, and applied Bonferroni corrections.

@@

We fit two full generalised linear models (GLM) with logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation using the proportion of environmental variation also secured in the prioritisation. The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using proportion of geographic variation also secured in the prioritisation. Both models also contained a variable indicating the species for which the prioritisations were generated, and an interaction term. They were subject to backwards step-wise term deletion routines to assess term significance. To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value (using the \texttt{pscl R} package; [@r492]).

Our third aim was to determine whether surrogate-based targets improved prioritisations. As previously mentioned, we generated prioritisations using different combinations of targets and conservation planning scenarios. We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations. Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets. We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" cominbation for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions. We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.

We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]). We fit a full model to the proportion of genetic variation secured in a given prioritisation. This model contained categorical variables indicating the combination of targets (amount-only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (single-species, multi-species, or multi-species with cost), the type of genetic variation measured (adaptive or neutral), and all interactions between them. Prioritisations generated for the same species were accomodated using a random intercept term. We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
